\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage[margin=2cm]{geometry}

\title{A Continuous Model Of Behavioural Phenotype: An Example Of
  Interaction Between \textit{Pieris Brasicae} and
  \textit{Trichogramma} Wasps}

\author{Kelly Black\thanks{Coresponding Author, kjblack@gmail.con},
  \and Malcolm R Adams, \and Aladeen Al Basheer, \and Caner Kazanci,
  \and Bernard Patten, \and Stuart Whipple, \and Sofya Zaytseva}

\begin{document}

\maketitle

\section{Introduction}

The variation in animal behaviours as well as the variation in
genetics for a given species plays a role in the broader population
dynamics of the species as well as the other animals that interact
with the species. A variety of mathematical models have been developed
to explore this important aspect of interactions. The models include
compartment models, stochastic, agent based models, as well as
combinations of these approaches. One difficulty with such approaches
is the lack of analytic tools to examine the general properties of the
resulting models.

We propose, instead, an approach assuming a continuous distribution
associated with one behavioural aspect of an animal's behaviour. The
example chosen is based on the interactions beteen parasitic wasps,
\textit{Trichogramma} Wasps, and the butterfly \textit{Pieris
  Brasicae} (cite Huigens).  The propensity of \textit{Pieris
  Brasicae} to employ a chemical associated with mating behaviour is
modeled as a continuous distribution, and a distributed model is
developed to approximate the resulting interactions.

In this treatment we approximate the system as consisting of two
species. The parastic wasp requires a relatively long interaction
time, and a predator-prey relationship is assumed. In particular, the
predation rates are approximated using a type II response. Due to the
variation in behaviour, though, we adapt the standard Holling response
function for this situation.

We first provide a more detailed overview of behavioural phenotypes
including some previous efforts to model the phenomena. Next, the
model used to provide insight into the phenomena is developed
including a re-derivation of a type II response assuming a continuous
variation in one behaviour. Following the derivation of the model, the
numerical approximation to the distributed system is briefly
discussed. Finally, some results from our numerical explorations are
provided.

\section{Behavioural Phenotype}

The primary motivation for why the distribution of how animals behave
and interact is given here. We begin with a discussion of the basic
idea of behavioural phenotype. Next, a brief overview of existing
modeling efforts is provided. Finally a specific example of social
spiders is discussed, and in this example the phenomena is directly
observed and provides context for the magnitude of the phenomena.

A number of authors have noted the importance of the role of variation
in a species' genetics or habits (cite Pinter-Wollman, Kortet,
Keiser). We begin this discussion with the survey provided by Kortet,
\textit{et al} (cite). The authors note that within any large group of
animals certain behaviours can vary across the population. They go on
to propose that the diversity of behaviours can profoundly impact the
broader dynamics of the population of the animals as well as animals
with a shared interdependence.

Kortet, \textit{et al} (cite) note the impact of variation in the
distribution of behaviour, they also note potential challenges this
recognition implies. One difficulty is identifying and quantifying the
distribution. An example they provide is the methods used to capture
individuals in a population can result in a biased estimate. The
reason for this is that some methods may be more or less likely to
capture a subset of individuals based on how bold or timid the animals
may be.

(Find example from their paper.)  Other mathematical models have been
developed.... Compartmental models by citation. Other agent based
models by citation. Hybrid models by citation.

Another, more detailed example of the importance of behavioural
differences between individuals in a population is provided by Keiser
\textit{et al} (cite). In their work, the differences in levels of
aggression in a social spider, \textit{Stegodyphus dumicola} are
examined. Experiments in sampling individuals found significant
differences in boldness levels. They also found that the variation of
boldness levels within a group impacted the spread of microbes across
the broader population.

A mathematical model for this system is provided by Pinter-Wollman,
\textit{et al} (cite). The focus of their work is in examining the
impact of different boldness levels of a key individual.  They note
that a small number of individuals within a colony may exhibit vastly
different levels of aggression, and a small number of individuals can
have a disproportionate impact in the broader success of the colony.

As a means to model this important aspect of a colony, Pinter-Wollman,
\textit{et al} provide a computational model consisting of an agent
based model with a stochastically generated population. Their approach
is based on a small number of rules that result in a distribution of
behaviours. The temporal interactions within the colony are
approximated via discrete time steps and stochastic decision
making. The external interactions are approximated using an
ordinary differential equation model.

In the resulting statistical analyses, Pinter-Wollman, \textit{et al}
found that different levels of boldness within key individuals
resulted in a difference in the mean boldness levels of a broader
group. For example, they found the differences in boldness impacted
the rate of capture of prey. They also found that the rules themselves
impacted also impacted the disease dynamics within the population.


\section{Modeling Genetic Variance As A Continuous Distribution}

Rather than construct a compartmental model or an agent based model, a
continuous, distributed parameter model is proposed here. The
resulting model for the example system is a coupled partial
differential equation and an ordinary differential equation. Prior to
the derivation itself an overview of the specific system, the
interactions between butterflies and a parasitic wasp, are
discussed. Next the model is derived. Finally, and analysis of a
simplified system of ordinary differential equations is examined as a
way to gain some preliminary insights into the resulting system.

\subsection{Butterflies v. Wasps}

The example of social spiders was briefly discussed earlier, but the
model developed here focuses on the interaction of butterflies,
\textit{Pieris Brasicae}, and \textit{Trichogramma} wasps that prey on
the butterflies' eggs. The interactions between these populations has
received a good deal of attention (cite many). We focus on the results
of one source in particular (cite Huigens). In the work presented by
(Huigens), the authors state that the pressure placed on the butterfly
population due to the wasps interaction results in a change in the
long term behaviour of the butterfly population.

The source of this pressure is the nature of the mating practices of
\textit{Pieris Brasicae}. In order to attract a mate, the female
butterflies tend to emit a pheromone designed to attract the males to
the females. After copulation, though, it is not in either the
female's nor the male's best interest to continue to attract other
butterflies. In response, the male butterflies have a propensity to
apply another pheromone, referred to as an anti-aphrodisiac, that will
dissuade reduce the effectiveness of the original pheromone emitted by
the female.

Some species of \textit{Trichogramma} wasps parasitize the eggs of the
butterflies. They do so by physically attaching themselves to the
butterfly and riding until the butterfly begins to attach their eggs
to a preferable plant. The wasps are able to detect the presence of
the anti-aphrodisiac and are able to recognize the increased
probability that a butterfly with the anti-aphrodisiac is more likely
to lay eggs. In response, the wasps are more likely to ride on a
butterfly if it detects the anti-aphrodisiac. It is important to note
this method of predation on a butterfly's eggs requires a non-trivial
time commitment with respect to handling and detecting the presence of
the eggs.

The authors (cite Huigens) note that the probability that a male will
make use of the anti-aphrodisiac varies between individuals. There are
different trade-offs in the use of the anti-aphrodisiac, but the
authors found that the wasps actions are applying direct pressure to
the habits of the butterflies. In particular they provide evidence
that the distribution of the propensity of the male butterflies to
make use of the anti-aphrodisiac is changing in time, and over time a
growing number of butterflies are less likely to engage in the
behaviour.

\subsection{Modeling Behavioural Phenotype As A Continuous
  Distribution}

\subsection{Stability Analysis Of The Simplified System}

\section{Numerical Approximation}

\section{Results}

\section{Conclusion}

\section{Acknowledgements}


\end{document}
